Commutative character sheaves and geometric types for supercuspidal representations

### 2017-07-04 Geometric types.tex
Working on Lemma 4.1

For connected, commutative groups, the standard method for geometrizing a character $\chi$ is to use the Lang isogeny $L : G \to G$, defined by $L(x) = x^{-1}\Frob{G}(x)$.  Then the pushforward $L_! (\EE)_G$ of the constant sheaf decomposes into a direct sum of rank $1$ local systems according to characters of the automorphism group $G(k)$ of the cover $L$.

@@

\begin{center}
\begin{spacing}{1.3}
\begin{tabular}{|l|l|}
\hline
$C_3$ & regular semisimple elements \\
$C_2$ & regular unipotent elements \\
$C_2^-$ & $\{-u : \mbox{ regular unipotent $u$}\}$ \\
$1$ & $\{1\}$ \\
$-1$ & $\{-1\}$ \\
\hline
\end{tabular}
\end{spacing}
\end{center}

%\vspace{0.1in}
\noindent
Each Weil character sheaf on $\SL_2$ over $\FFF$ is the perverse extension of one of the following local systems: the rank-$1$ constant sheaf $\mathbf{1}_{C_3}$ (resp. $\mathbf{1}_{C_2}$, $\mathbf{1}_{C_2^-}$, $\mathbf{1}_{1}$,  $\mathbf{1}_{-1}$) on $C_3$ (resp. $C_2$, $C_2^-$, $1$, $-1$), or the non-trivial local system $\cs{E}_{C_2}$  (resp.  $\cs{E}_{C_2^-}$) trivialized by the  double cover of $C_2$ (resp. $C_2^-$).
The following table describes these perverse sheaves by showing how they decompose in the Grothendieck group into conjugation-equivariant standard sheaves on $\SL_2$, and also gives the values of the trace of Frobenius of each Weil character sheaf on $\SL_2$ on the $7$ conjugacy classes in $\SL_2(\FFF)$.

@@

\begin{spacing}{1.3}
\[
\begin{array}{|l|l|ccccccc|}
\hline
\text{Perverse} & \text{Standard} & \multicolumn{7}{|c|}{\text{Trace of Frobenius values}} \\ \cline{3-9}
\text{Sheaf} & \text{Sheaves} & (\begin{smallmatrix}1 & 0\\ 0 & 1\end{smallmatrix}) & (\begin{smallmatrix}2 & 0\\ 0 & 2\end{smallmatrix}) & (\begin{smallmatrix}0 & 1\\ 2 & 0\end{smallmatrix})  & (\begin{smallmatrix}1 & 1\\ 0 & 1\end{smallmatrix})  & (\begin{smallmatrix}1 & 2\\ 0 & 1\end{smallmatrix}) & (\begin{smallmatrix}2 & 1\\ 0 & 2\end{smallmatrix})  & (\begin{smallmatrix}2 & 2\\ 0 & 2\end{smallmatrix}) \\
\hline
\IC(\mathbf{1}_{C_3})  &  \mathbf{1}_{C_3}[3]  & -1 & -1 & -1 & -1 & -1 & -1 & -1 \\
 \IC(\mathbf{1}_{C_2})  &  \mathbf{1}_{C_2}[2] \oplus \mathbf{1}_1[0]  & 1 & 0 & 0 & 1 & 1 & 0 & 0 \\
 \IC(\cs{E}_{C_2})  &  \cs{E}_{C_2}[2]  & 0 & 0 & 0 &  \sqrt{-3}  &  -\sqrt{-3} & 0 & 0 \\
 \IC(\mathbf{1}_{C_2^-})  &  \mathbf{1}_{C_2^-}[2] \oplus \mathbf{1}_{-1}[0]  & 0 & 1 & 0 & 0 & 0 & 1 & 1 \\
 \IC(\cs{E}_{C_2^-})  &  \cs{E}_{C_2^-}[2]  & 0 & 0 & 0 & 0 & 0 &   \sqrt{-3}  &   -\sqrt{-3}  \\
 \IC(\mathbf{1}_1)  &  \mathbf{1}_1[0]  & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
 \IC(\mathbf{1}_{-1})  &  \mathbf{1}_{-1}[0]  & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
\hline
\end{array}
\]
\end{spacing}

@@

\noindent 
Using the trace of Frobenius values, we find that 
\[
- \IC(\mathbf{1}_{C_3}) - \frac32 \IC(\mathbf{1}_{C_2}) + \frac12 \IC(\cs{E}_{C_2}) - \frac32 \IC(\mathbf{1}_{C_2^-}) - \frac12 \IC(\cs{E}_{C_2^-}) + \frac32 \IC(\mathbf{1}_1) + \frac32 \IC(\mathbf{1}_{-1})
\]
also geometrizes $\chi$.  Note that all seven character sheaves appear in this geometrization, even though we start just with a character.

@@

\begin{table}[ht]
\caption{Notation conversion chart.}
\begin{spacing}{1.3}
\begin{tabular}{| c|l | l | }
\hline
\text{this paper} & \text{Jiu-Kang Yu, {\it Construction of tame}} & \cite{yu:01a} \\
 & \text{{\it supercuspidal representations}} &  \\
\hline
$\oK^0$ & $\,^\circ K^0 = G^0(F)_y$ & \cite{yu:01a}*{\S 15} \\
$\oK^{i+1}$ & $\,^\circ K^{i+1} = (\,^\circ K^0) \vec{G}^{(i+1)}(F)_{y,(0, s_0, \ldots, s_{i})}$ & \cite{yu:01a}*{\S 15} \\
$\orho^0$ & $\,^\circ \rho^0$ & \cite{yu:01a}*{\S 15} \\
$\orho^{i+1}$ & $\,^\circ \rho^{i+1}$ &  \cite{yu:01a}*{\S 15} \\
$\varphi^i$ & $\phi_i\vert_{\,^\circ K^i }$ & \cite{yu:01a}*{\S 3}\\
$J_{i+1}$ & $J^{i+1} = (G^i,G^{i+1})(F)_{y, (r_i, s_i)}$ & \cite{yu:01a}*{\S 3} \\
$V_{i+1}$ & $J^{i+1}/J^{i+1}_+ = (G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ (G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}$ & \cite{yu:01a}*{\S 3} \\ 
$V_{i+1}^\sharp$ & $(G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ \ker(\widehat{\phi}_i\vert_{(G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}})$ &  \cite{yu:01a}*{\S 4} \\
$Z_{i+1}$ & $\ker(V_{i+1}^\sharp\to V_{i+1})$ & \cite{yu:01a}*{\S 11} \\
$(f_{j+1}, j_{i+1})$ & $(f,j)$ & \cite{yu:01a}*{\S 11} \\
$\langle\ ,\ \rangle_{i+1}$ & $\langle\ ,\ \rangle$ & \cite{yu:01a}*{\S 11} \\
\hline
\end{tabular}
\end{spacing}
\label{table:notation}
\end{table}%

>>>>

For connected, commutative groups, the standard method for geometrizing a character $\chi$ is to use the Lang isogeny $L : G \to G$, defined by $L(x) = x^{-1}\Frob{G}(x)$.  The pushforward $L_! (\EE)_G$ of the constant sheaf decomposes into a direct sum of rank $1$ local systems according to characters of the automorphism group $G(k)$ of the cover $L$.

Now drop the assumption that $G$ is commutative, but assume that $G(k) \subset \image(L)$ (which will hold if $G$ is connected).  We apply a similar strategy to construct a local system on $G$ whose trace of Frobenius is $\chi$.

\begin{lemma}
If $\chi \in G(k)^*$ then there is a character $\chi' \in G(k)^*$ so that $\chi'(yxy^{-1}) = \chi(x)$ for all $x \in G(k)$ and $y \in L^{-1}(x^{-1})$.
\end{lemma}
\begin{proof}
The condition $y \in L^{-1}(x^{-1})$ is equivalent to $\Frob{}(y) = yx^{-1}$, and thus $yxy^{-1} \in G(k)$ since $\Frob{}(yxy^{-1}) = (yx^{-1})x(xy^{-1}) = yxy^{-1}.$
Given $z \in G(k)$, we want to write $z = yxy^{-1} = y\Frob{}(y^{-1}) = L(y^{-1})$ in order to define $\chi'(z) = \chi(x)$.  This is possible by our assumption that $G(k) \subset \image(L)$.  To see that $\chi'$ is well-defined, suppose that $z = y_1 x_1 y_1^{-1} = y_2 x_2 y_2^{-1}$ with $y_i \in L^{-1}(x_i^{-1})$.  Then $y_1 \Frob{}(y_1^{-1}) = y_2 \Frob{}(y_2^{-1})$ so $y_1^{-1} y_2 \in G(k)$ and $y_2^{-1} y_1 \in G(k)$.  We can now write
\begin{align*}
x_1 x_2^{-1} &= y_1^{-1} z y_1 y_2^{-1} z^{-1} y_2 \\
&= (y_1^{-1} y_2) (y_2^{-1} z y_2) (y_2^{-1} y_1) (y_2^{-1} z^{-1} y_2) \in G(k)\der.
\end{align*}
Thus $\chi(x_1 x_2^{-1}) = 1$ so $\chi(x_1) = \chi(x_2)$ and $\chi'$ is well defined. \todo{Finish proof here}
\end{proof}

@@

\begin{tabular}{|l|l|}
\hline
$C_3$ & regular semisimple elements \\
$C_2$ & regular unipotent elements \\
$C_2^-$ & $\{-u : \mbox{ regular unipotent $u$}\}$ \\
$1$ & $\{1\}$ \\
$-1$ & $\{-1\}$ \\
\hline
\end{tabular}

@@

Suppose that $C$ is a stratum and $x \in C$. Local systems on $C$ are classified up to isomorphism by irreducible representations of $\pi_0(Z_G(x))$.  This component group is trivial for $C_3$, $1$ and $-1$, and of order $2$ for $C_2$ and $C_2^-$.  Denote the nontrivial local systems on $C_2$ and $C_2^-$ by $\cs{E}_{C_2}$ and $\cs{E}_{C_2^-}$.  Then the character sheaves on $\SL_2$ are the following, where the second column gives a decomposition in the Grothendieck group of standard sheaves on $G$.

@@

\begin{tabular}{|l|l|ccccccc|}
\hline
Perverse & Standard & \multicolumn{7}{|c|}{Trace of Frobenius values} \\ \cline{3-9}

Sheaf & Sheaves & \smat1001 & \smat2002 & \smat0120 & \smat1101 & \smat1201 & \smat2102 & \smat2202 \\
\hline
$\IC(\mathbf{1}_{C_3})$ & $\mathbf{1}_{C_3}[3]$ & -1 & -1 & -1 & -1 & -1 & -1 & -1 \\
$\IC(\mathbf{1}_{C_2})$ & $\mathbf{1}_{C_2}[2] \oplus \mathbf{1}_1[0]$ & 1 & 0 & 0 & 1 & 1 & 0 & 0 \\
$\IC(\cs{E}_{C_2})$ & $\cs{E}_{C_2}[2]$ & 0 & 0 & 0 & $\sqrt{-3}$ & $-\sqrt{-3}$ & 0 & 0 \\
$\IC(\mathbf{1}_{C_2^-})$ & $\mathbf{1}_{C_2^-}[2] \oplus \mathbf{1}_{-1}[0]$ & 0 & 1 & 0 & 0 & 0 & 1 & 1 \\
$\IC(\cs{E}_{C_2^-})$ & $\cs{E}_{C_2^-}[2]$ & 0 & 0 & 0 & 0 & 0 & $\sqrt{-3}$ & $-\sqrt{-3}$ \\
$\IC(\mathbf{1}_1)$ & $\mathbf{1}_1[0]$ & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
$\IC(\mathbf{1}_{-1})$ & $\mathbf{1}_{-1}[0]$ & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
\hline
\end{tabular}

\vspace{0.1in}
\noindent Using the trace of Frobenius values, we find that 
\[
- \IC(\mathbf{1}_{C_3}) - \frac32 \IC(\mathbf{1}_{C_2}) + \frac12 \IC(\cs{E}_{C_2}) - \frac32 \IC(\mathbf{1}_{C_2^-}) - \frac12 \IC(\cs{E}_{C_2^-}) + \frac32 \IC(\mathbf{1}_1) + \frac32 \IC(\mathbf{1}_{-1})
\]
also geometrizes $\chi$.  Note that all seven character sheaves appear in this geometrization, even though we start just with a character.

@@

\begin{table}[ht]
\caption{Notation conversion chart.}
\begin{spacing}{1.2}
\begin{tabular}{| c|l | l | }
\hline
\text{this paper} & \text{Jiu-Kang Yu, {\it Construction of tame}} & \cite{yu:01a} \\
 & \text{{\it supercuspidal representations}} &  \\
\hline
$\oK^0$ & $\,^\circ K^0 = G^0(F)_y$ & \cite{yu:01a}*{\S 15} \\
$\oK^{i+1}$ & $\,^\circ K^{i+1} = (\,^\circ K^0) \vec{G}^{(i+1)}(F)_{y,(0, s_0, \ldots, s_{i})}$ & \cite{yu:01a}*{\S 15} \\
$\orho^0$ & $\,^\circ \rho^0$ & \cite{yu:01a}*{\S 15} \\
$\orho^{i+1}$ & $\,^\circ \rho^{i+1}$ &  \cite{yu:01a}*{\S 15} \\
$\varphi^i$ & $\phi_i\vert_{\,^\circ K^i }$ & \cite{yu:01a}*{\S 3}\\
$J_{i+1}$ & $J^{i+1} = (G^i,G^{i+1})(F)_{y, (r_i, s_i)}$ & \cite{yu:01a}*{\S 3} \\
$V_{i+1}$ & $J^{i+1}/J^{i+1}_+ = (G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ (G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}$ & \cite{yu:01a}*{\S 3} \\ 
$V_{i+1}^\sharp$ & $(G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ \ker(\widehat{\phi}_i\vert_{(G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}})$ &  \cite{yu:01a}*{\S 4} \\
$Z_{i+1}$ & $\ker(V_{i+1}^\sharp\to V_{i+1})$ & \cite{yu:01a}*{\S 11} \\
$(f_{j+1}, j_{i+1})$ & $(f,j)$ & \cite{yu:01a}*{\S 11} \\
$\langle\ ,\ \rangle_{i+1}$ & $\langle\ ,\ \rangle$ & \cite{yu:01a}*{\S 11} \\
\hline
\end{tabular}
\end{spacing}
\label{table:notation}
\end{table}%
